---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r}
library(alluvial)
library(plyr)
library(rjson)

api_url <- 'https://journaltransfer.issn.org/api?query=*'
download.file(api_url, './transferlist.json')
dat <- fromJSON(file = 'transferlist.json')

# clean the data to something usable
tidy_dat <- data.frame()
for (i in 1:length(dat)) {
  tidy_dat <- rbind(tidy_dat, data.frame(id = dat[[i]]$id,
                             origin = dat[[i]]$contents$transPubName,
                             dest = dat[[i]]$contents$recPubName))
}

red_dat <- ddply(tidy_dat, .(origin, dest), function (x) {
  return(data.frame(x, freq = dim(x)[1]))
})

red_dat <- red_dat[,2:4]
red_dat <- unique(red_dat)

red_dat$origin <- as.character(red_dat$origin)
red_dat$dest <- as.character(red_dat$dest)

red_dat <- red_dat[order(-red_dat$freq),]
```

# Tweet 1 

i had a free afternoon and decided to do this, here's some stats

of all `r length(tidy_dat$id)` documented transfers on the issn transfer page, there is concentration of journals (not distribution) in who owns the journals: `r length(unique(red_dat$origin))` original publishers transferred the `r length(tidy_dat$id)` journals to `r length(unique(red_dat$dest))` receiving pubs

# Tweet 2 

```{r}
oligarchs <- sort(table(red_dat$dest), decreasing = TRUE)[1:10]
perc <- (sum(oligarchs)/1605) * 100
```
the top ten out of `r length(unique(red_dat$dest))` receiving publishers got `r round(perc, 0)`% of the original journals: `r names(oligarchs)`

# tweet 3

notable one who isn't there: Elsevier. It looks like they clearly didn't care for acquisitions, which befits the data based business model they're more busy with (see https://twitter.com/chartgerink/status/1190201681432268801?s=20)

# tweet 5

and then of course, let's do a nice visual overview to wrap that up. I highlight only the top ten receiving publishers, otherwise it becomes a bit much

```{r fig.cap="Flow of journals from one to another publisher", fig.asp=.625}
nams <- names(oligarchs)
cols <- ifelse(red_dat$dest == nams[1], 'purple', 
                      ifelse(red_dat$dest == nams[2], 'blue', 
                             ifelse(red_dat$dest == nams[3], 'magenta', 
                                    ifelse(red_dat$dest == nams[4], 'black', 'grey'))))
alluvial(red_dat[order(red_dat$dest),1:2], freq = red_dat$freq,
         col = cols,
         border = cols,
         hide = red_dat$freq <= min(oligarchs),
         cex = 0.01)
```
